---
title: "Solution for the assignment of the eight class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(broom)
library(psych)
library(Hmisc)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

# 1. PCA on Diener flourishing scale

First, we select the variables of interest.

```{r}
diener_data <- dplyr::select(processed, contains("diener"))
```

In R, we can run the PCA with just one function. We can set the scale argument to true to scale our variables to unit variance and center the variables.

```{r}
pca <- 
  diener_data %>% 
  prcomp(scale = TRUE, center = TRUE)

summary(pca)
```

As the data are standardized the standard deviation values for each principal component are the Eigenvalues.

As we can see only the first PC has an Eigenvalue greater than 1.

The first PC accounts for 71.02% of the variance, while the second PC accounts for the 6.13% of the variance.

Together they explain 77.15% of the variance.

Lets plot the results.

```{r}
pca %>%
  tidy(matrix = "eigenvalues") %>% 
  ggplot(aes(PC, percent)) +
  geom_col(fill = "#56B4E9", alpha = 0.8) +
  scale_x_continuous(breaks = 1:8) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  ) +
  theme_minimal()
```

# 2. PCA of MET

First, we select the variables of interest.

```{r}
met_subscales <- c("jollet", "savor", "a_vhat", "onreg", "rezil")

met_data <- dplyr::select(processed, all_of(met_subscales))
```

We run the PCA while standardizing the variables.

```{r}
pca_met <- 
  met_data %>% 
  prcomp(scale = TRUE, center = TRUE)

summary(pca_met)
```

As we can see only the first PC has an Eigenvalue greater than 1, and it accounts for 59.54% of the variance.

```{r}
pca_met %>%
  tidy(matrix = "eigenvalues") %>% 
  ggplot(aes(PC, percent)) +
  geom_col(fill = "#56B4E9", alpha = 0.8) +
  scale_x_continuous(breaks = 1:5) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  ) +
  theme_minimal()
```

# 3. Use varimax and promax rotation on the first three principal components of the second task.

I could use the `prcomp` function ad before, but we would have to use a cutting point for the Eigenvalues to only include 3 PC. Thus, I will use the `psych::principal` function, where I can set the number of PC in the function call.

Thus, I run a new PCA with three components and varimax rotation. This function returns the standardized scores by default. The function calculates the correlation matrix by pairwise deletion of the missing values.

```{r}
pca_met_varimax <- psych::principal(met_data, rotate = "varimax", nfactors = 3, scores = TRUE)
```

Lets take a look at the Eigenvalues.

```{r}
pca_met_varimax$value
```

The rest of the results.

```{r}
pca_met_varimax
```

The factor loadings are quite small for _onreg_ and _rezil_ on RC1. Based on the communalities (h2) it seems that most of the variance explained by these two items can be explained by other items.

The complexity scores suggest that _jollet_ and maybe _a_vhat_ the items measure more than one latent construct.

The proportion var suggest that the RC1 explains 38% of the variance in the responses, while RC3 27% and RC2 21%.

The fit based upon off diagonal values support a good fit.

Now we should run the same analysis but allow factors to correlate, therefore we use an oblique rotation called promax.

```{r}
pca_met_promax <- psych::principal(met_data, rotate = "promax", nfactors = 3, scores = TRUE)
```

Lets see the results of this solution.

```{r}
pca_met_promax
```

Allowing for correlation reduced the complexity and we have a much cleaner latent structure than before with varimax. RC1 highly correlates (0.66) with RC3 and only medium wiht RC2 (0.29), there is a medium to high correlation (0.42) between RC3 and RC2.

# 4. Save the factors from the promax rotation and correlate the with the PERMA test's items. Which items has the highest correlattion with the first factor.

Calculating the factor scores.

```{r}
scores <- factor.scores(met_data, pca_met_promax)

scores <- as.data.frame(scores$scores)
```

Binding calculated scores to the original data.

```{r}
perma_vars <- c("p_poz_erz", "p_elmely", "p_poz_kapc", "p_ert_cel", "p_telj", "p_boldog", "p_egeszs", "p_neg_erz", "p_magany")

processed_scores <- 
  processed %>% 
  bind_cols(., scores) %>% 
  select(RC1, RC2, RC3, all_of(perma_vars))
```

Running the correlations.

```{r}
corr_matrix <- rcorr(as.matrix(processed_scores))

corr_matrix$r
```

Interestingly there is a strong correlation between almost all of the variables.The strongest correlation is between RC1 and and the PERMA items is with _p_poz_erz_ (r = 0.69)