---
title: "Solution for the assignment of the eleventh class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(broom)
library(psych)
library(lavaan)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

# 1. Run CFA on the best model from the sixths task in the ninth assignments best model.

Get the data for the model.

```{r}
vars <- c("p_elmeny_percent", "testi_fi", "alt_lelki", "alt_eg_all", "fizero", "arcocska", "aggodalo", "ideges", "feszult", "nyugtala")

task_data <- 
  processed %>% 
  select(all_of(vars))
```

First I will run the EFA as I did not run it in the _ninth_assignment_.

With 2 factors.

```{r}
efa2 <- fa(task_data, nfactors = 2, rotate = "varimax", fm = "ml")

efa2
```

With 3 factors.

```{r}
efa3 <- fa(task_data, nfactors = 3, rotate = "varimax", fm = "ml")

efa3
```

The BIC is smaller for the 2 factor model so I choose that one. However, three items have a high complexity and they load for both factors.

Specifying the model for CFA.

```{r}
model <- '
factor1 =~ p_elmeny_percent + arcocska + aggodalo + ideges + feszult + nyugtala
factor2 =~ testi_fi + alt_lelki + fizero  + alt_eg_all'
```

Fitting the model.

```{r}
cfa <- cfa(model, data = task_data)

summary(cfa, standardized = TRUE, fit.measures = TRUE)
```

# 2. How good is the model fit?

The model has a significantly good fit with _X^2(34, 500)_ = 305.43, _p_ < 0.001.

# 3. What about the other test statistics?

The RMSEA = 0.129 CI90[0.116, 0.142] with _p_ < 0.001 for alpha <= 0.05.

# 4. What about CFI, SRMR, and TLI?

The Standardized Root Mean Square Residual (SRMR) is 0.129. The value ranges between 0 and 1, and as it is higher than the usually accepted 0.08 cutoff value, it does not indicate a good fit.

The TLI has a penalty for adding more parameters. It is 0.893 which is below the .9 cutoff point for a good model.

Tha CFI is usually highly correlated with the TLI. The value here is 0.919 which means it is not good either.

# 5. Create a graph for the latent structure.

Plotting the model with standardized model parameter estimates.

```{r}
semPlot::semPaths(cfa, "std")
```

# 6. Which variables should be dropped?

I suspect that if I drop the two variables with high complexity the model will be better.

Lets test it for fun!

```{r}
task_data_filtered <-
  task_data %>% 
  select(-alt_lelki, arcocska)

model_filtered <- '
factor1 =~ p_elmeny_percent + aggodalo + ideges + feszult + nyugtala
factor2 =~ testi_fi + fizero  + alt_eg_all'

cfa_filtered <- cfa(model_filtered, data = task_data_filtered)

summary(cfa_filtered, standardized = TRUE, fit.measures = TRUE)
```

Indeed the CFI and TLI values are higher!