---
title: "Solution for the assignment of the fourth class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(psych)
library(kableExtra)
library(ez)
library(WRS2)
library(MBESS)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

_The first three tasks are the same tasks as in assignment three. However, here, I supplement the analysis with confidence intervals and figures of the means._

# 1. On which scales do people differ the most in their education level?

To investigate this question I decided to run several one-way ANOVA tests with the education level as the independent variable, and the different scales as the dependent variable. I will include only interval scales. Then, I will compare the differences with a standardized effect size measure, eta-square. The comparison with the biggest eta-square will be the answer.

Also, I will run the ANOVA tests in bulk in one iteration as I did in the second assignment. This way I can save some time typing, which is always nice. The results will be saved in a table format for an easy comparison. Also, a plot with the different standardized effect sizes would be helpful.

I am running type 2 ANOVA tests.

```{r, warning=FALSE, message=FALSE}
# Selecting and saving interval scales
interval_vars <- 
  processed %>% 
  select(26:50) %>% 
  names()

# Converting participant id and isk to factor variable
isk_scale_data <-
  processed %>% 
  mutate(isk = as.factor(isk),
         index = as.factor(index))

# Checking the order of levels for isk as a factor variable
levels(isk_scale_data$isk)

# Function to run the one-way ANOVA
isk_anova <- function(data, dv_var) {
  eval(
    substitute(
      # The ANOVA
      ezANOVA(data = data, dv = dv_var, wid = index, between = isk, detailed = TRUE),
      list(dv_var = dv_var)
      )
    )
}

isk_means <- function(var) {
  isk_scale_data %>% 
    group_by(isk) %>% 
    summarise(mean = mean(.data[[var]]),
              sd = sd(.data[[var]]),
              n = n(),
              se = sd / sqrt(n),
              ci_lower = mean - 1.96 * se,
              ci_upper = mean + 1.96 * se)
}

# Running the ANOVA tests
isk_scale_res <- 
  tibble(
    variable = interval_vars,
    anova_res =  map(variable,
                     ~ isk_anova(data = isk_scale_data, dv_var = .x)),
    f_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "F")),
    dfn = map_dbl(anova_res, ~ pluck(.x, 2, "DFn")),
    dfd = map_dbl(anova_res, ~ pluck(.x, 2, "DFd")),
    p_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "p")),
    generalized_eta_square = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "ges")),
    means = map(variable, isk_means)
    ) %>% 
  mutate(
    ci = MBESS::ci.pvaf(F.value = f_value, df.1 = dfn, df.2 = dfd, N = 500, conf.level=.95)
    ) %>% 
  arrange(desc(generalized_eta_square))

isk_scale_res %>% 
  select(-anova_res, -means, -dfd, -dfn) %>%
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared", "CI"),
    align = c("l", "c", "c", "c","c"),
    caption = "One-way ANOVA Results for Each Scale Variable by Three Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

The figures show the each education level on the X axis and the corresponding scale on the Y axis. The errobars are the 95% confidence intervals.

```{r}
apa_pointplot <- function(data, scale_name) {
  data %>%
    unnest() %>% 
    ggplot2::ggplot() +
    ggplot2::aes(x = isk, y = mean) +
    ggplot2::geom_point(size = 4) +
    ggplot2::geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = .5) +
    ggplot2::labs(y = scale_name) +
    papaja::theme_apa()
}

map2(isk_scale_res$means, isk_scale_res$variable, ~ apa_pointplot(.x, .y))
```

To answer the question I only consider significant effects. The highest difference between the educational levels were present on the _avhat_ variable which is a subscale of the MET test (@andras_jollet_2020).

Lets see the results on a plot as well.

```{r}
isk_scale_res %>% 
  mutate(significant = case_when(p_value <= 0.05 ~ "Significant",
                                 p_value > 0.05 ~ "Nonsignificant"),
         variable = fct_reorder(variable, generalized_eta_square)) %>% 
  ggplot() +
  aes(x = variable, y = generalized_eta_square, shape = significant) +
  geom_point(size = 3) +
  coord_flip() +
  labs(
    y = "Eta-squared",
    x = "Scales",
    shape = "Significance"
  ) +
  papaja::theme_apa() +
  theme(
    legend.position = c(.9, .35)
  )
```

# 2. On which scales do people differ more if we merge the data of the two lowest educational level?

```{r}
# Merge the levels
isk_scale_merged_data <- 
  processed %>% 
  mutate(isk = case_when(isk == "altalanos" ~ "altalanos_es_kozepiskola",
                         isk == "kozepiskola" ~ "altalanos_es_kozepiskola",
                         TRUE ~ isk),
         isk = as.factor(isk))

# Running the ANOVA tests
isk_scale_merged_res <- 
  tibble(
    variable = interval_vars,
    anova_res =  map(variable,
                     ~ isk_anova(data = isk_scale_merged_data, dv_var = .x)),
    f_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "F")),
    dfn = map_dbl(anova_res, ~ pluck(.x, 2, "DFn")),
    dfd = map_dbl(anova_res, ~ pluck(.x, 2, "DFd")),
    p_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "p")),
    generalized_eta_square = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "ges")),
    means = map(variable, isk_means)
    ) %>% 
  mutate(
    ci = MBESS::ci.pvaf(F.value = f_value, df.1 = dfn, df.2 = dfd, N = 500, conf.level=.95)
    ) %>% 
  arrange(desc(generalized_eta_square))

isk_scale_merged_res %>% 
  select(-anova_res, -means, -dfd, -dfn) %>%
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared", "CI"),
    align = c("l", "c", "c", "c", "c"),
    caption = "One-way ANOVA Results for Each Scale Variable by Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

The figures show the each education level on the X axis and the corresponding scale on the Y axis. The errobars are the 95% confidence intervals. Interestingly, the CIs for the 'altalanos' educational level are wide as the N is low for this group.

```{r}
apa_pointplot <- function(data, scale_name) {
  data %>%
    unnest() %>% 
    ggplot2::ggplot() +
    ggplot2::aes(x = isk, y = mean) +
    ggplot2::geom_point(size = 4) +
    ggplot2::geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = .5) +
    ggplot2::labs(y = scale_name) +
    papaja::theme_apa()
}

map2(isk_scale_merged_data$means, isk_scale_merged_data$variable, ~ apa_pointplot(.x, .y))
```

# 3. Run a two-way ANOVA with educational level and gender on the scale type variables. Is there a scale where the interaction significant? How big is the eta-squared here?

```{r}
# Converting neme to a factor variable
isk_scale_merged_data <- 
  isk_scale_merged_data %>% 
  mutate(neme = as.factor(neme),
         index = as.factor(index))

# Function to run the two-way ANOVA
isk_neme_anova <- function(data, dv_var) {
  eval(
    substitute(
      # The ANOVA
      ezANOVA(data = data, dv = dv_var, wid = index, between = .(isk, neme), detailed = TRUE),
      list(dv_var = dv_var)
      )
    )
}

# Running the ANOVA tests
isk_neme_scale_res <- 
  tibble(
    variable = interval_vars,
    anova_res = map(variable,
                     ~ isk_neme_anova(data = isk_scale_merged_data, dv_var = .x)[["ANOVA"]]),
    interaction_res = map(anova_res, . %>% filter(Effect == "isk:neme")),
    f_value = map_dbl(interaction_res, . %>% pull(F)),
    p_value = map_dbl(interaction_res, . %>% pull(p)),
    generalized_eta_square = map_dbl(interaction_res, . %>% pull(ges))
    ) %>% 
  arrange(desc(generalized_eta_square))

isk_neme_scale_res %>% 
  select(-anova_res, -interaction_res) %>%
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "Results of the Interaction Terms of the Two-way ANOVA for Each Scale Variable by Education Level and Gender"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

# 4. Is there a difference between the means of the subscales of the Globalis Jollet scale?

To answer this question first I select the needed variables and trasnform them to long format.

```{r}
global_comparison_data <-
  processed %>% 
  select(index, g_jpszi, g_jszoc, g_jspir, g_jerzpsz) %>% 
  gather(key = "subscale", value = "value", -index)
```

_The means and SDs of each subscale._

```{r}
global_comparison_data %>% 
  group_by(subscale) %>% 
  summarise(mean = mean(value),
            sd = sd(value),
            n = n(),
            se = sd / sqrt(n),
            ci_lower = mean - 1.96 * se,
            ci_upper = mean + 1.96 * se)
```

Running the ANOVA.

```{r}
ezANOVA(dv = value, wid = index, within = subscale, data = global_comparison_data)
```

According to the one-way repeated measures ANOVA there is a significant difference between the subscales.

_Therefore, I can now run posthoc analyses. I use the Bonferroni correction to control the family-wise error rates._

```{r}
pairwise.t.test(global_comparison_data$value, global_comparison_data$subscale, p.adj = "bonf")
```

_The results of the pairwise pot-hoc comparisons shows that only "g_jszoc " differed significantly from all of the other subscales._

# 5. Two-way ANOVA on the five subscales of MET with gender and educational level as between subject factors, and control for age and financial state as covariants.

The data transformation for this analysis was done previously by converting gender and educational level to factor variables and merging the two lowest N levels of the educational level.

The ezANOVA function automatically controls for the covariates supplied in the _between_covariates_ argument.

When interpreting the results we have to take into account the group sizes are unbalanced.

```{r}
# Function to run the two-way ANOVA
isk_neme_cov_anova <- function(data, dv_var) {
  eval(
    substitute(
      # The ANOVA
      ezANOVA(data = data, dv = dv_var, wid = index, between = .(isk, neme), between_covariates = .(eletkora, anyagi) ,detailed = TRUE),
      list(dv_var = dv_var)
      )
    )
}

met_subscales <- c("jollet", "savor", "a_vhat", "onreg", "rezil")

# Running the ANOVA tests
isk_neme_anova_cov_res <- 
  tibble(
    variable = met_subscales,
    anova_res = map(variable,
                     ~ isk_neme_cov_anova(data = isk_scale_merged_data, dv_var = .x)[["ANOVA"]]),
    interaction_res = map(anova_res, . %>% filter(Effect == "isk:neme")),
    f_value = map_dbl(interaction_res, . %>% pull(F)),
    p_value = map_dbl(interaction_res, . %>% pull(p)),
    generalized_eta_square = map_dbl(interaction_res, . %>% pull(ges))
    ) %>% 
  arrange(desc(generalized_eta_square))

isk_neme_anova_cov_res %>% 
  select(-anova_res, -interaction_res) %>%
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "Results of the Interaction Terms of the Two-way ANOVA for Each Scale Variable by Education Level and Gender and by Controlling for Age and Financial Level Covariants"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

According to the results there is no significant interaction between gender and the educational level on none of the subscales.

Lets look at the same results when we did not control for age and financial situation.

```{r}
isk_neme_scale_res %>% 
  dplyr::filter(variable %in% met_subscales) %>% 
  select(-anova_res, -interaction_res)
```

It seems like that none of the interactions were significant even before controlling for the age and financial situation.