---
title: "Solution for the assignment of the second class"
author: "Kovacs Marton"
date: "9/21/2021"
output:
  pdf_document:
    latex_engine: xelatex
bibliography: multistat_references.bib
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(psych)
library(kableExtra)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
raw <- read_tsv("data/boldog.txt", locale = locale(encoding = "UTF-8"))
```

R sometimes have a problem with special characters. I will fix that now.

```{r}
# first I fix the special Hungarian characters
Encoding(names(raw)) <- 'latin1'
# I translate them to English letters for the sake of simplicity
iconv(names(raw), from = "latin1", to = "ASCII//TRANSLIT")
```

Also, I like to use _snake_case_ for my variable names, therefore I will transform the variable names.

```{r}
raw <- janitor::clean_names(raw)
```

# Data exploration

```{r}
skimr::skim(raw) %>% 
  kable()
```

# Transforming variables

```{r}
processed <- 
  raw %>% 
  mutate(neme = case_when(neme == 1 ~ "ferfi",
                          neme == 2 ~ "no", 
                          TRUE ~ NA_character_),
         isk = case_when(isk == 1 ~ "altalanos",
                         isk == 2 ~ "kozepiskola",
                         isk == 3 ~ "foiskola",
                         isk == 4 ~ "egyetem",
                         TRUE ~ NA_character_))
```

# 1. Which variables are intervalum scale variables?

Based on the name of the variables and on a paper focusing on the validation of the MET test (@andras_jollet_2020), I decided that I would consider the following variables are on an interval scale:

```{r, results = "asis"}
# select variables
interval_vars <- 
  processed %>% 
  select(eletkora, jollet, 26:50) %>% 
  names()

# print as a bullet list
cat(paste('-', interval_vars), sep = '\n')
```
  
# 2. & 3. Which of these variables are the least looking like a normal distribution, and which of these are normally distributed?

Plotting each variable.

```{r, message=FALSE, warning=FALSE}
map(interval_vars, ~ apa_hist(processed, .x))
```

Looking at the skewness and the kurtosis values. I will use 1.5 and -1.5 as arbitrary cutoff values for indicating a non normal distribution.

```{r}
tibble(
  variable = interval_vars,
  skewness = map_dbl(variable, ~ skew(processed[[.x]])),
  kurtosis = map_dbl(variable, ~ kurtosi(processed[[.x]])),
  non_normal = case_when(skewness > 1.5 | skewness < -1.5 | kurtosis > 1.5 | kurtosis < -1.5 ~ "not normal",
                         TRUE ~ "normal")
  ) %>% 
  kable(
    format = "latex",
    # booktabs = TRUE,
    # escape = FALSE,
    col.names = c("Variable", "Skewness", "Kurtosis", "Normally distributed"),
    align = c("l", "c", "c", "c"),
    caption = "Investigation of the Skewness and Kurtosis Values of the interval Variables"
  ) %>% 
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE)
```

Based on visual investigation and the kurtosis and skewness values the _eletkora_  variables seems like the most normally distributed variable. Whereas, the _p_ert_cel_ and the _p_telj_ variables are not normally distributed according to the cutoff values.

# 4. Running Kolmogorov-Smirnov tests

I am running two-sided tests with a normal distribution as the cumulative distribution function.

```{r, warning=FALSE}
tibble(
  variable = interval_vars,
  kolmogorov_res = map(interval_vars, ~ ks.test(processed[[.x]], y = "pnorm")),
  test_statistic = map_chr(kolmogorov_res, "statistic"),
  p = map_dbl(kolmogorov_res, "p.value")
  ) %>% 
  mutate(p = if_else(p == 0, "p < 0.01", as.character(p))) %>% 
  select(variable, test_statistic, p) %>%
    kable(
    format = "latex",
    # booktabs = TRUE,
    # escape = FALSE,
    col.names = c("Variable", "Test statistic", "p"),
    align = c("l", "c", "c"),
    caption = "Results of the Kolmogorov-Smirnov Tests"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE)
```

All tests are significant.

# 5. On which scales do we find the largest difference between male and female? What are the corresponding standardised effectsizes (in Cohens' d and eta square)?

For all of the variables I used the mean of the variable by gender group for the comparison. Also, I am running type 2 ANOVAs to calculate the eta square from.

```{r}
# creating a function for calculating the mean difference
gender_mean_diff <- function(var) {
  processed %>% 
    group_by(neme) %>% 
    summarise(mean = mean(.data[[var]], na.rm = TRUE)) %>% 
    spread(neme, mean) %>% 
    mutate(diff = no - ferfi) %>% 
    pull(diff)
}

# creating function to calculate pooled SD
gender_sd_pooled <- function(var) {
  processed %>% 
    group_by(neme) %>% 
    summarise(sd = sd(.data[[var]], na.rm = TRUE)) %>% 
    spread(neme, sd) %>% 
    mutate(sd_pooled = sqrt((ferfi^2 + no^2)/2)) %>% 
    pull(sd_pooled)
}


tibble(
  variable = interval_vars,
  mean_diff = map_dbl(variable, ~ gender_mean_diff(.x)),
  sd_pooled = map_dbl(variable, ~ gender_sd_pooled(.x)),
  cohens_d = mean_diff / sd_pooled,
  anova_res = map(variable, 
                  ~ aov(processed[[.x]] ~ processed[["neme"]], data = processed)),
  eta_squared = map_dbl(anova_res, ~ lsr::etaSquared(.x, type = 2, anova = FALSE)[[1]])
  ) %>% 
  select(variable, mean_diff, cohens_d, eta_squared) %>%
    kable(
    format = "latex",
    # booktabs = TRUE,
    # escape = FALSE,
    col.names = c("Variable", "Mean difference", "Cohens' d", "Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "The Corresponding Raw and Standardised Effect Sizes For Each intervalum Variable"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE)
```

# References {-}

<div id="refs"></div>