---
title: "Solution for the assignment of the seventh class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(Hmisc)
library(regclass)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

# 1. Look for a variable thats variance is explained by the 8 Diener items by at least 65%.

I will first calculate the total score for the Diener flourishing scale.

```{r}
diener_data <-
  processed %>% 
  mutate(diener_sum = diener1 + diener2 + diener3 + diener4 + diener5 + diener6 + diener7 + diener8)
```

For this task I am investigating only interval variables. To calculate the Rsquared I calculate the Pearson correlation coefficient and square it.

```{r}
interval_vars <- 
  processed %>% 
  select(26:50) %>% 
  names()

rsq <- function (data, x, y) cor(data[[x]], data[[y]]) ^ 2

diener_res <-
  tibble::tibble(
    variable = interval_vars,
    rsquared = map_dbl(variable,
                  ~ rsq(diener_data, .x, "diener_sum"))
    ) %>% 
  mutate(
    rsquared = round(rsquared, 2),
    flag = case_when(rsquared >= .65 ~ TRUE,
                     rsquared < .65 ~ FALSE)
  ) %>% 
  arrange(desc(rsquared))
```

The total score on the Diener 8 item scale only explains more than 65% of the variance of the _g_jerzpsz_ scale.

# 2. Lets test whether adding aggodalom and ideges variables as main effects will increase the R2.

To do this we have to create a linear regression model.

```{r}
m <- lm(g_jerzpsz ~ diener_sum + aggodalo + ideges, data = diener_data)

summary(m)
```

Adding those two variables indeed increased the Rsquared to 0.6991.

# 3. Transforming the kor variable

```{r}
kor_data <-
  processed %>% 
  dplyr::select(eletkora, index) %>% 
  # bind_cols(kor_data, poly(kor_data$eletkora), degree = 3)
  mutate(korz = scale(eletkora),
         korz2 = korz^2,
         korz3 = korz^3,
         korz4 = korz^4)
```

Now, lets create a correlation matrix including these variables.

```{r}
Hmisc::rcorr(as.matrix(kor_data), type = "pearson")
```

# 4. Predicting PERMA by korz with polynomial regression. Which power has the largest effect on the outcome variable? Plot the results.

```{r}
polyreg_data <-
  kor_data %>% 
  left_join(., select(processed, perma, index), by = "index")

m <- lm(perma ~ korz + korz2 + korz3 + korz4, data = polyreg_data)

summary(m)
```

The second and fourth degree polynomial variables predicted the perma outcome variable significantly.

# 5. Look for a variable that has a third degree polynomial relationsip with age. Plot the relationship.

To look for the relationship I will look at the correlation between the interval variables and the third degree polynomial of age. I am looking for a significant relationship.

```{r}
third_data <-
  processed %>% 
  left_join(., kor_data, by = "index")

third_res <-
  tibble::tibble(
    variable = interval_vars,
    cor_res = map(variable, 
                  ~ my_cor(
                    data = third_data,
                    x = .x,
                    y = "korz3",
                    method = "spearman"
                  )),
    cor_r = map_dbl(cor_res, ~ pluck(.x, "estimate", "rho")),
    cor_p = map_dbl(cor_res, ~ pluck(.x, "p.value")),
    flagged = case_when(cor_p <= 0.05 ~ TRUE,
                        TRUE ~ FALSE)
  ) %>% 
  arrange(desc(cor_r))
```

The results show that 10 variables had a significant third degree polynomial relationship with the age variable. The following table is in descending order based on the Spearmans' rho.

```{r}
third_res %>% 
  select(-cor_res)
```

Now, I would like to create scatterplot showing the relationship between these variables and age.

```{r}
third_plot <- 
  third_res %>% 
  dplyr::filter(flagged) %>% 
  mutate(plot = map(variable, 
                    ~ poly_plot(
                      data = third_data,
                      x = "korz3",
                      y = .x
                    ))
         )

third_plot$plot
```

# 6. Multiple linear regression where pelmeny is the outcome variable and Aggodalom, Ideges, Feszült, Nyugtalan are the predictor variables

```{r}
m <- lm(p_elmeny_percent ~ aggodalo + ideges + feszult + nyugtala, data = processed)

summary(m)
```

The rsquared suggest that the predictor variables explain 25.85% of the variance in the outcome variable. 

For the predictor variables the VIF is the highest for the _feszult_ variable.

```{r}
vif <- VIF(m)

vif
```

And the tolerances are:

```{r}
1 / vif
```

Plotting the standardized residuals.

```{r}
to_plot <- broom::augment(m)

ggplot(to_plot,
       aes(x = .fitted, y = .resid)) +
  geom_point() +
  papaja::theme_apa()
```
