---
title: "Solution for the assignment of the sixth class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(psych)
library(kableExtra)
library(ez)
library(WRS2)
library(GPArotation)
library(MBESS)
library(rstatix)
library(ARTool)
library(FSA)
library(ddst)
library(Hmisc)
library(ltm)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

# 1. What are the correlations and rank correlations between Globalis Jollet Scale and MET?

First I create a table for all the combinations between the variables.

```{r}
global <- dplyr::select(processed, 31:36) %>% names()
met <- dplyr::select(processed, 25:29) %>% names
correlation_res <- tidyr::crossing(global, met)
```

The I calculate the parametric correlation for each combination with a Pearson correlation, and the rank correlation with the Spearman method.

```{r}
correlation_res <-
  correlation_res %>% 
  mutate(
    pearson_res = map2(global, met, 
                     ~ my_cor(
                        data = processed,
                        x = .x,
                        y = .y,
                        method = "pearson")),
    pearson_r = map_dbl(pearson_res, ~ pluck(.x, "estimate", "cor")),
    pearson_p = map_dbl(pearson_res, ~ pluck(.x, "p.value")),
    spearman_res = map2(global, met, 
                     ~ my_cor(
                        data = processed,
                        x = .x,
                        y = .y,
                        method = "spearman")),
    spearman_r = map_dbl(spearman_res, ~ pluck(.x, "estimate", "rho")),
    spearman_p = map_dbl(spearman_res, ~ pluck(.x, "p.value")),
  )

correlation_res %>% 
  dplyr::select(-pearson_res, -spearman_res) %>% 
  arrange(desc(pearson_r))
```

The highest correlation was between _gjerz_ and the _jollet_ scales.

# 2. Find scales that are in a non-linear relationship with any of the subscales of MET.

# 3. What is the proportion of the negative and positive relationships between the resiliance scale and the total score of the Diener scale.

I am not sure that I understand the question correctly, but I will calculate the correlation between the _resil_ variable and each item of the Diener flourishing scale. Than, I will look at the proportion of the positive and negative correlations based on the Pearson correlation coefficient.

```{r}
diener <- dplyr::select(processed, contains("diener")) %>% names()
rezil <- c("rezil")

diener_rezil_res <- tidyr::crossing(rezil, diener)

diener_rezil_res <-
  diener_rezil_res %>% 
  mutate(
    res = map2(rezil, diener, 
                     ~ my_cor(
                        data = processed,
                        x = .x,
                        y = .y,
                        method = "pearson")),
    r = map_dbl(res, ~ pluck(.x, "estimate", "cor")),
    r = round(r, 2),
    valence = case_when(r > 0 ~ "positive",
                        r < 0 ~ "negative")
  ) %>% 
  dplyr::select(-res) %>% 
  arrange(desc(r))

diener_rezil_res %>% 
  count(valence) %>% 
  mutate(prop = n / 8 * 100)
```

# 4. How big is the Cronbachs alpha for the total score of the Diener scale?

```{r}
diener_data <-
  processed %>% 
  dplyr::select(contains("diener"))

cronbach.alpha(diener_data, CI = TRUE)
```

# 5. Same for each gender group.

```{r}
diener_male_data <-
  processed %>% 
  filter(neme == "ferfi") %>% 
  dplyr::select(contains("diener"))

cronbach.alpha(diener_male_data, CI = TRUE)

diener_female_data <-
  processed %>% 
  filter(neme == "no") %>% 
  dplyr::select(contains("diener"))

cronbach.alpha(diener_female_data, CI = TRUE)
```

# 6. How big is the Mcdonalds' omega for the Diener flourishing scale?

```{r}
psych::omega(diener_data, nfactors = 1)
```

# 7. How big is the Cronbachs alpha and Mcdonalds' omega for the scale consisting of the Testi_fi, Áltlelki, Áltegáll, Fizerö, Arcocska, Aggodalo, Ideges, Feszült, Nyugtalan items?

Some of the items are on reversed scales so we have to reverse them. To run the analyses I also have to drop missing items.

```{r}
scale_data <-
  processed %>% 
  dplyr::select(8:16) %>% 
  filter_all(all_vars(!is.na(.)))

# Looking at the maximum values for each scale
scale_data %>% 
  summarise_all(max, na.rm = TRUE)

# I only suspect that if these items belong to the same scale the maximum value for all of them is 7
reverse <- function(x) {
    7 - x
}

scale_data <-
  scale_data %>% 
  mutate_at(vars(arcocska, aggodalo, ideges, feszult, nyugtala), reverse)
```

We can now calculate the Cronbachs alpha and Mcdonalds' omega. For calculating the omega, I propose that the items are grouped into one factor.

```{r}
cronbach.alpha(scale_data, CI = TRUE)

psych::omega(scale_data, nfactors = 1)
```
