---
title: "Solution for the assignment of the tenth class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(broom)
library(psych)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
processed <- read_tsv("data/boldog_processed.tsv")
```

# Data exploration

```{r}
skimr::skim(processed) %>% 
  kable()
```

# 1. Run PCA with varimax rotation

Select the variables.

```{r}
vars <- c("p_elmeny_percent", "testi_fi", "alt_lelki", "alt_eg_all", "fizero", "arcocska", "aggodalo", "ideges", "feszult", "nyugtala")

task_data <- 
  processed %>% 
  select(all_of(vars))
```

Create a correlation matrix for the scree plot.

```{r}
cor_matrix <- cor(task_data, use = "pairwise.complete.obs")

cor_matrix
```

Make the scree plot.

```{r}
scree(cor_matrix)
```

We have two PCs with bigger than 1 Eigenvalues.

Run the PCA.

```{r}
pca <- psych::principal(task_data, rotate = "varimax", nfactors = 2, scores = TRUE)

pca
```

Based on the results the two PCs explain 71% of the variance. Not bad!

# 2. Run EFA on the same variables

```{r}
efa <- fa(task_data, nfactors = 2, rotate = "varimax", fm = "pf")

efa 
```

With varimax rotation looking for 2 factors, and with PAF estimation (due to the non normal distributions of the variables), I do not find any variables with lower communality than 0.25 (see h2).

# 3. Check out KMO and Bartletts test results!

```{r}
KMO(cor_matrix)
```

The overall KMO score and the individual items scores are all good.

```{r}
cortest.bartlett(cor_matrix, n = 500, diag = FALSE)
```

The p value is so small that the test rounds it to 0. We can run an EFA the variables are good for factor analysis.

# 4. Is the model bad based on significance?

```{r}
efa
```

The RMSEA is 0.093 CI90[0.079, 0109].

_The total number of observations was  500  with Likelihood Chi Square =  139.67  with prob <  1.6e-17_

# 5. Use a promax rotation

```{r}
efa_promax <- fa(task_data, nfactors = 2, rotate = "promax", fm = "pf")

efa_promax
```

p_elmeny_percent, arcocska, aggodalo, ideges, feszult, nyugtala loads for the 1st factor, but with p_elmeny_percent loading negatively. alt_lelki and arcocska has a high complexity. 

Here it is possible for the items to correlate with each other.
