---
title: "Solution for the assignment of the third class"
author: "Kovacs Marton"
date: "9/29/2021"
output:
  pdf_document:
    latex_engine: xelatex
bibliography: multistat_references.bib
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = TRUE)
# loading necesarry packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(psych)
library(kableExtra)
library(ez)
library(WRS2)
#loading custom R functions
r_scripts <- list.files("R/", full.names = TRUE)
walk(r_scripts, source)
```

# Importing data

```{r}
raw <- read_tsv("data/boldog.txt", locale = locale(encoding = "UTF-8"))
```

R sometimes have a problem with special characters. I will fix that now.

```{r}
# first I fix the special Hungarian characters
Encoding(names(raw)) <- 'latin1'
# I translate them to English letters for the sake of simplicity
iconv(names(raw), from = "latin1", to = "ASCII//TRANSLIT")
```

Also, I like to use _snake_case_ for my variable names, therefore I will transform the variable names.

```{r}
raw <- janitor::clean_names(raw)
```

I will use this dataset for further assignments, so I save the cleaned version in .tsv format.

```{r}
write_tsv(raw, "data/boldog_raw.tsv")
```

# Data exploration

```{r}
skimr::skim(raw) %>% 
  kable()
```

# Transforming variables

```{r}
processed <- 
  raw %>% 
  mutate(neme = case_when(neme == 1 ~ "ferfi",
                          neme == 2 ~ "no", 
                          TRUE ~ NA_character_),
         isk = case_when(isk == 1 ~ "altalanos",
                         isk == 2 ~ "kozepiskola",
                         isk == 3 ~ "foiskola",
                         isk == 4 ~ "egyetem",
                         TRUE ~ NA_character_))
```
I will also save this version with the more informative labels for future use.

```{r}
write_tsv(processed, "data/boldog_processed.tsv")
```

# 1. On which scales do people differ the most in their education level?

To investigate this question I decided to run several one-way ANOVA tests with the education level as the independent variable, and the different scales as the dependent variable. I will include only interval scales. Then, I will compare the differences with a standardized effect size measure, eta-square. The comparison with the biggest eta-square will be the answer.

Also, I will run the ANOVA tests in bulk in one iteration as I did in the second assignment. This way I can save some time typing, which is always nice. The results will be saved in a table format for an easy comparison. Also, a plot with the different standardized effect sizes would be helpful.

I am running type 2 ANOVA tests.

```{r, warning=FALSE, message=FALSE}
# Selecting and saving interval scales
interval_vars <- 
  processed %>% 
  select(26:50) %>% 
  names()

# Converting participant id and isk to factor variable
isk_scale_data <-
  processed %>% 
  mutate(isk = as.factor(isk),
         index = as.factor(index))

# Checking the order of levels for isk as a factor variable
levels(processed$isk)

# Function to run the one-way ANOVA
isk_anova <- function(data, dv_var) {
  eval(
    substitute(
      # The ANOVA
      ezANOVA(data = data, dv = dv_var, wid = index, between = isk, detailed = TRUE),
      list(dv_var = dv_var)
      )
    )
}

# Running the ANOVA tests
isk_scale_res <- 
  tibble(
    variable = interval_vars,
    anova_res =  map(variable,
                     ~ isk_anova(data = isk_scale_data, dv_var = .x)),
    f_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "F")),
    p_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "p")),
    generalized_eta_square = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "ges"))
    ) %>% 
  select(-anova_res) %>% 
  arrange(desc(generalized_eta_square))

isk_scale_res %>% 
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "One-way ANOVA Results for Each Scale Variable by Three Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

To answer the question I only consider significant effects. The highest difference between the educational levels were present on the _avhat_ variable which is a subscale of the MET test (@andras_jollet_2020).

Lets see the results on a plot as well.

```{r}
isk_scale_res %>% 
  mutate(significant = case_when(p_value <= 0.05 ~ "Significant",
                                 p_value > 0.05 ~ "Nonsignificant"),
         variable = fct_reorder(variable, generalized_eta_square)) %>% 
  ggplot() +
  aes(x = variable, y = generalized_eta_square, shape = significant) +
  geom_point(size = 3) +
  coord_flip() +
  labs(
    y = "Eta-squared",
    x = "Scales",
    shape = "Significance"
  ) +
  papaja::theme_apa() +
  theme(
    legend.position = c(.9, .35)
  )
```

# 2. On which scales do people differ more if we merge the data of the two lowest educational level?

```{r}
# Merge the levels
isk_scale_merged_data <- 
  processed %>% 
  mutate(isk = case_when(isk == "altalanos" ~ "altalanos_es_kozepiskola",
                         isk == "kozepiskola" ~ "altalanos_es_kozepiskola",
                         TRUE ~ isk),
         isk = as.factor(isk))

# Running the ANOVA tests
isk_scale_merged_res <- 
  tibble(
    variable = interval_vars,
    anova_res =  map(variable,
                     ~ isk_anova(data = isk_scale_merged_data, dv_var = .x)),
    f_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "F")),
    p_value = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "p")),
    generalized_eta_square = map_dbl(anova_res, ~ pluck(.x, "ANOVA", "ges"))
    ) %>% 
  select(-anova_res) %>% 
  arrange(desc(generalized_eta_square))

isk_scale_merged_res %>% 
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "One-way ANOVA Results for Each Scale Variable by Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

# Run a two-way ANOVA with educational level and gender on the scale type variables. Is there a scale where the interaction significant? How big is the eta-squared here?

```{r}
# Converting neme to a factor variable
isk_scale_merged_data <- 
  isk_scale_merged_data %>% 
  mutate(neme = as.factor(neme),
         index = as.factor(index))

# Function to run the two-way ANOVA
isk_neme_anova <- function(data, dv_var) {
  eval(
    substitute(
      # The ANOVA
      ezANOVA(data = data, dv = dv_var, wid = index, between = .(isk, neme), detailed = TRUE),
      list(dv_var = dv_var)
      )
    )
}

# Running the ANOVA tests
isk_neme_scale_res <- 
  tibble(
    variable = interval_vars,
    anova_res = map(variable,
                     ~ isk_neme_anova(data = isk_scale_merged_data, dv_var = .x)[["ANOVA"]]),
    interaction_res = map(anova_res, . %>% filter(Effect == "isk:neme")),
    f_value = map_dbl(interaction_res, . %>% pull(F)),
    p_value = map_dbl(interaction_res, . %>% pull(p)),
    generalized_eta_square = map_dbl(interaction_res, . %>% pull(ges))
    ) %>% 
  select(-anova_res, -interaction_res) %>%
  arrange(desc(generalized_eta_square))

isk_neme_scale_res %>% 
  kable(
    format = "latex",
    col.names = c("Variable", "F value", "p value", "Generalized Eta-squared"),
    align = c("l", "c", "c", "c"),
    caption = "Results of the Interaction Terms of the Two-way ANOVA for Each Scale Variable by Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE) %>% 
  add_footnote("Results are arranged in descending order by the size of the effec size.")
```

# 4. Run the same as before but with a robust equivalent of the ANOVA.

As a robust non-parametric alternative for the two-way ANOVA I will use the medians instead of the means for the calculation.

```{r}
# Function to run the robust two-way ANOVA
isk_neme_robust_anova <- function(dv_var) {
  pbad2way(as.formula(paste(dv_var, "~ isk:neme")), data = isk_scale_merged_data, est = "median")
}

# Running the ANOVA tests
isk_neme_scale_robust_res <- 
  tibble(
    variable = interval_vars,
    anova_res = map(variable,
                     ~ isk_neme_robust_anova(dv_var = .x)),
    p_value = map_dbl(anova_res, ~pluck(.x, "AB.p.value"))
    ) %>% 
  select(-anova_res) %>% 
  mutate(significance = case_when(p_value <= 0.05 ~ "Significant",
                                 p_value > 0.05 ~ "Nonsignificant"))

isk_neme_scale_robust_res %>% 
  kable(
    format = "latex",
    col.names = c("Variable", "p value", "Significance"),
    align = c("l", "c", "c"),
    caption = "Results of the Interaction Terms of the Robust Two-way ANOVA for Each Scale Variable by Education Level"
  ) %>%
  row_spec(row = 0, align = "c") %>%
  kable_styling(full_width = TRUE)
```

# 5. Is there a difference between the means of the subscales of the Globalis Jollet scale?

To answer this question first I select the needed variables and transform them to long format.

```{r}
global_comparison_data <-
  processed %>% 
  select(index, g_jpszi, g_jszoc, g_jspir, g_jerzpsz) %>% 
  gather(key = "subscale", value = "value", -index)

# Running the ANOVA
ezANOVA(dv = value, wid = index, within = subscale, data = global_comparison_data)
```

According to the one-way repeated measures ANOVA there is a significant difference between the subscales.

# 6. Two-way mixed ANOVA with the four subscales and gender.

Again, we start with the data transformation.

```{r}
global_comparison_gender_data <-
  processed %>% 
  select(index, neme, g_jpszi, g_jszoc, g_jspir, g_jerzpsz) %>% 
  gather(key = "subscale", value = "value", -index, -neme)

# Running the ANOVA
ezANOVA(dv = value, wid = index, within = subscale, between = neme, data = global_comparison_gender_data)
```

According to the results there is a significant interaction between the subscales and the gender. Also, the subscales have a significant main effect, but not gender.

# References {-}

<div id="refs"></div>